---
title: "SC19093 neuron - astrocyte co-cultures"
subtitle: "Integrate individual sample into one Seurat data object"
author: Stephanie Strohbuecker
output: html_notebook
---

The following QC and pre-processing follows the workflow described in the [Seurat v3 tutorial](https://satijalab.org/seurat/v3.0/pbmc3k_tutorial.html).

```{r}
RMD_file_name <- "1-1.Integrate_samples"
```


## Libraries

```{r}
library(tidyverse)
library(here)

library(ggpubr)

library(Seurat)
```
Enable parallelization with the `future` package as descibed [here](https://satijalab.org/seurat/v3.0/future_vignette.html)
```{r}
library(future)
plan("multiprocess", workers = 4)
```

```{r load_data}
# Obtain the metadata information for each of the samples
meta_data <- openxlsx::read.xlsx(
  here(
    "input_data",
    "SC19093_ASF_extended_sample_description.xlsx"
  ),
  startRow = 3,
  colNames = TRUE
) %>%
  dplyr::slice(-(1:2)) %>%
  dplyr::slice(-NROW(.)) %>%
  select_if(~ !all(is.na(.))) %>%
  filter(!is.na(sample_description)) %>%
  dplyr::select(
    Sample = `Sample/LIMSID`,
    Name = `Sample/Name`,
    sample_description,
    sample_id = short_name
  ) %>%
  mutate(sample_id_t = sample_id) %>%
  separate(sample_id_t,
           into = c("cell_type", "treatment"),
           sep = "_")

# File path to the base Cellranger count output
file_path_10x <- here("input_data", "Cellranger")

wang_marker <- openxlsx::read.xlsx(
  here(
    "input_data",
    "marker_gene_lists",
    "wang_et_al_2018_Science",
    "DER-21_Single_cell_markergenes_UMI.xlsx"
  ),
  startRow = 2
)

# Path to output and figures directories
output <- str_c(here(), "output", RMD_file_name, sep = "/")

if (!dir.exists(output)) {
  dir.create(output, recursive = TRUE)
}

figures_out <- str_c(here(), "figures", RMD_file_name, sep = "/")

if (!dir.exists(figures_out)) {
  dir.create(figures_out, recursive = TRUE)
}

```

```{r}
sample_path_10x <-
  dir(
    file_path_10x,
    recursive = TRUE,
    pattern = "filtered_feature_bc_matrix$",
    include.dirs = TRUE,
    full.names = TRUE
  )

sample_path_10x <- set_names(sample_path_10x, str_extract(sample_path_10x, "GAN343A[0-9]*"))
```


# Read data and generate the Seurat object

```{r}
SC19093_seurat <- map2(sample_path_10x, names(sample_path_10x), function(file_path, sample_name){
  proj_name <- meta_data %>% filter(Sample == sample_name) %>% pull(sample_id)
  proj_name
  data_10x <- Read10X(file_path)
  CreateSeuratObject(counts = data_10x, project = proj_name,
                     min.cells = 0, min.features = 0) # instead of the initial filtering done in script 02.QC_individual_samples.Rmd, remove filtering parameters to ensure that all Seurat objects contain the same number of features
})
SC19093_seurat
```


```{r change_list_names}
SC19093_seurat <- SC19093_seurat %>% 
  set_names(map_chr(names(SC19093_seurat), ~meta_data %>% 
                      filter(Sample == .x) %>% 
                      pull(sample_id)))

SC19093_seurat <- SC19093_seurat[order(names(SC19093_seurat))]
```

```{r}
SC19093_seurat <- map(SC19093_seurat, function(Seur_obj){
  Seur_obj[["percent_mito"]] <- PercentageFeatureSet(Seur_obj, pattern = "^MT-")
  Seur_obj
})
```

```{r}
# Filter low quality cells based on the cut-offs identified in script 02.QC_individual_samples.Rmd
SC19093_seurat_filtered <- map(SC19093_seurat, function(Seur_obj) {
  subset(Seur_obj, subset = nFeature_RNA > 200 &
           nFeature_RNA < 3000 & percent_mito < 10)
})
```

```{r}
# Normalize and find variable features in each of the individual datasets
SC19093_seurat_filtered <-
  map(SC19093_seurat_filtered, function(Seur_obj) {
    NormalizeData(Seur_obj)
  })
SC19093_seurat_filtered <-
  map(SC19093_seurat_filtered, function(Seur_obj) {
    FindVariableFeatures(Seur_obj,
                         selection.method = "vst",
                         nfeatures = 2000)
  })
```


```{r}
# Identify anchors for subsequent data integration, the dimensionality of the dataset is set to 30 by default bu may be varied over a broad range, e.g. between 10 and 50
anchors <- FindIntegrationAnchors(object.list = SC19093_seurat_filtered, dims = 1:30)
```
Integrate the data using the anchors identified in the previous step. For the subsequent regression of the effect of the cell cycle set the `features.to.integrate` to all features, i.e. not just the variable ones as described [here](https://github.com/satijalab/seurat/issues/1500).
```{r}
data_integrated <-
  IntegrateData(
    anchorset = anchors,
    dims = 1:30,
    features.to.integrate = rownames(SC19093_seurat_filtered$astro_ctrl)
  )
```

```{r}
# Change the default assay from the "RNA" (raw data) to the "integrated" one; note that variable features of this assay are automatically set during IntegrateData
DefaultAssay(data_integrated) <- "integrated"
```

```{r}
# Seurat cell cycle genes
s_genes <- cc.genes$s.genes
g2m_genes <- cc.genes$g2m.genes
```

Based on the expression of the above G2M/M and S phase markes, each cell is assigned a score. These marker sets should be anticorrelated in their expression levels, and cells expressing neither are likely not cycling and in G1 phase.
Seurat does not use the discrete classifications (G2M/G1/S) in downstream cell cycle regression ad uses instead the quantitative scores for G2M and S phasee.

```{r}
# Calculate cell cycle scores
data_integrated <-
  CellCycleScoring(
    data_integrated,
    s.features = s_genes,
    g2m.features = g2m_genes,
    set.ident = TRUE
  )
```

```{r eval=FALSE, include=FALSE}
# regress out the cell cycle effect; in this form this code chunk takes a very long time implement with future package (https://satijalab.org/seurat/v3.0/future_vignette.html)
data_integrated <-
  ScaleData(
    data_integrated,
    vars.to.regress = c("S.Score", "G2M.Score"),
    features = rownames(data_integrated)
  )
```

```{r eval=FALSE, include=FALSE}
saveRDS(
  data_integrated,
  file = str_c(
    output,
    "Seurat_integrated_datasets_cell_cycle_regressed.rds",
    sep = "/"
  )
)
```

```{r}
data_integrated <- readRDS( str_c(
    output,
    "Seurat_integrated_datasets_cell_cycle_regressed.rds",
    sep = "/"
  ))
```


```{r}
data_integrated <- RunPCA(data_integrated, npcs = 100, ndims.print = 1:5, nfeatures.print = 5)
```

```{r}
data_integrated <-
  JackStraw(data_integrated, num.replicate = 100, dims = 100)
data_integrated <- ScoreJackStraw(data_integrated, dims = 1:100)
```

```{r, fig.width=12}
JackStrawPlot(data_integrated, dims = 1:75)
```


```{r}
ElbowPlot(data_integrated, ndims = 100)
```

```{r fig.height=12}
DimHeatmap(data_integrated, dims = c(1:3, 43:51), cells = 500, balanced = TRUE)
```

We selected a dimensionality of 50 for the subsequent clustering.

```{r}
data_integrated <- FindNeighbors(data_integrated, reduction = "pca", dims = 1:50)
```

```{r}
data_integrated <- FindClusters(data_integrated, resolution = c(0.4, 0.6, 0.8, 1.0, 1.2))
```

```{r}
data_integrated <- RunUMAP(data_integrated, reduction = "pca", dims = 1:50)
```

```{r}
data_integrated@meta.data <- data_integrated[[]] %>% separate(old.ident, into = c("cell_type", "treatment"), sep = "_")
```

```{r}
DefaultAssay(data_integrated) <- "RNA"
```

```{r}
data_integrated <- ScaleData(data_integrated, features = rownames(data_integrated))
```

```{r}
DefaultAssay(data_integrated) <- "integrated"
```


```{r eval=FALSE, include=FALSE}
saveRDS(
  data_integrated,
  file = str_c(
    output,
    "Seurat_integrated_datasets_cell_cycle_regressed_clustered_UMAP.rds",
    sep = "/"
  )
)
```



