---
title: "Pre-processing scRNA-Seq data for deconvolution using Scaden"

author: Karishma D'Sa
output: 
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    toc_float: yes
    toc_depth: 6
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---


# Scaden
## Background

## Data/file formats
<b>File format for the sc data:</b><br>
  1. Cell type labels file :  should be of size (n x 1), where n is the number of cells in the data. <br>
   The single column in this file has to be labeled 'Celltype'. 
   Extra columns may be present, as long as the 'Celltype' column which specifies the cell type labels in the correct order is present.
<br><br>

  2. Counts file : count data should be of size (n x g), where g is the number of genes and n is the number of samples. The order must be the same as for the cell type labels.

<br><br>

Scaden requires the normalised or raw counts of the scRNA-Seq data (https://scaden.readthedocs.io/en/latest/usage.html#file-formats). <br>

<br>
Generating the above files for the 6 samples <br>
astrocytes - treated and untreated <br>
neurons - treated and untreated <br>
coculture - treated and untreated <br><br>
<br>

For the clustering, using resolution 0.25<br>


```{r rmdName, include = F}
RMD_file_name <- "2-2.NormalizeData_4Scaden"
```

```{r libraries,  include = F}
suppressPackageStartupMessages(library(Seurat))
library(here)

library(tidyverse)
library(here)

library(ggpubr)


library(DT)
library(data.table)
#Enable parallelization with the `future` package as descibed #(https://satijalab.org/seurat/v3.0/future_vignette.html)
library(future)
plan("multiprocess", workers = 4)

```

```{r output path, include = F }
# Path to output and figures directories
output <- str_c(here(), "output", RMD_file_name, sep = "/")

if (!dir.exists(output)) {
  dir.create(output, recursive = TRUE)
}

```


# Using the integrated seurat object- with additional clustering

Rawcounts of all samples from the clustering resolution 0.25 combined in a single file <br>
Also generated a normalised counts file
8132 cells ; 22588 genes


```{r getting the clustering as a single file, eval = F, include = F}
clustering_res <- 0.25
res_prefix <- "integrated_snn_res."


counts_df <- as.data.frame( t( as.matrix( GetAssayData(object = data_integrated, assay = "RNA", slot = "counts"))))

counts_df <- counts_df[ ,colSums(counts_df[])>0]  #excluding genes with 0 counts in all cells, left with 22588
cat("\t cCells=", nrow(counts_df)) #8132 cells
cat("\t nGenes=", length(colnames(counts_df)))
  
cell_types <- data_integrated[[paste0(res_prefix, clustering_res) ]] # 8132 cells
head(cell_types, 2)
names(cell_types) <- "Celltype"

stopifnot( identical (rownames(cell_types), rownames(counts_df)))  #checking the order of the cells matches

cell_types <- data.frame(lapply(cell_types, as.character), stringsAsFactors=FALSE)

cell_types$Celltype <- paste0("celltype_", cell_types$Celltype )
  
rownames(counts_df) <- NULL
rownames(cell_types) <- NULL

  #saving
write_tsv(counts_df, paste0(output, "/raw_counts/ipsc_allsamples_raw_counts.txt"), quote=F)
write_tsv(cell_types, paste0(output, "/raw_counts/ipsc_allsamples_celltypes.txt"), quote=F)

```

