---
title: "Investigating the cell type proportions in the bulk, post deconvolution"
author: "Karishma D'Sa"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_float: yes
    toc_depth: 6
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---

#Aim 
To see if there is a significant difference in the cell type proportions in the treated vs untreated astrocytes and co-cultures. 


```{r libraries, echo=F, include =F}
library(here)
library(data.table)
library(ggplot2)
library(DT)
library(knitr)

source(here("./R/common_sample.R"))
source(here("./R/plot_setting.R"))
```

```{r load the data, echo=F}
celltype_propfile <-here("../results/deconv_scaden/allsamples_run/predictions_nsamples/scaden_pred_comb_var0_1_cells3000_samples2000.txt") 

ct_prop <- fread(celltype_propfile)
setnames(ct_prop, old ="V1", new="sample")

sInfo <- get_sample_info()
sInfo$sample <- paste0(sInfo$cells, sInfo$treatment, "_", sInfo$indi)
#head(ct_prop, 3); head(sInfo, 2);length(unique(sInfo$sample))
sInfo <- unique( sInfo[, .(sample, cells, treatment, indi)] )
ct_prop <- merge(ct_prop, sInfo, by = "sample", all.x=T)
ct_prop[ treatment== "1uM_syn_oligomer", treatment := "treated"]
ct_prop[ treatment== "no_treatment", treatment := "untreated"]
ct_prop[ cells== "astrocyte_neuron", cells := "co-culture"]

unique(ct_prop$cells)
rm(sInfo)
```

#Results

From the scRNA-seq data, celltypes 0 and 4 are likely to be astrocytic and the remaining neuronal. <br>
<br>

##Table of cell type proportions{.tabset .tabset-fade .tabset-pills}

Celltype proportion*100

```{r table of the cell type proportions, include =F}
ct_prop_copy <- copy(ct_prop)
names(ct_prop_copy) <- gsub("celltype_" , "ct_", names(ct_prop_copy))
cols_upd <- names(ct_prop_copy)[grep("ct_", names(ct_prop_copy))] 


ct_prop_copy[, (cols_upd) := lapply(.SD, function(x) 
        format(x*100,  digits = 2 )), .SDcols = cols_upd]

data_source <- unique(ct_prop_copy$cells)
out <- vector(mode = "character", length = length(data_source))
for (i in data_source) {
  out[i] <- knit_expand(text = c("### {{stringr::str_to_title(i)}}",
                                 "```{r, echo = FALSE}",
                                 "ct_prop_copy[cells == '{{i}}',.(treatment, indi, ct_0,ct_1,ct_2,ct_3,ct_4,ct_5,ct_6,ct_7 )]",
                                 "```"))
}

```

`r paste(knit(text = out), collapse = '\n')`

## Avg cell type proportion per cell culture and treatment
Avg cell type proportion*100
```{r table of the mean cell type proportions, include =F}

head(ct_prop_copy, 2)
#cols_upd <- names(ct_prop_copy)[grep("ct_", names(ct_prop_copy))]
ct_prop_copy$cells_treatment <- paste0(ct_prop_copy$cells, "_", ct_prop_copy$treatment)
#ct_prop_copy[, (cols_upd) := lapply(.SD, mean), .SDcols = cols_upd, by = cells_treatment ]
ct_prop_copy[, paste0("avg_", (cols_upd)) :=lapply(.SD, function(x) 
        mean(as.numeric(x),  digits = 2 )),  by = cells_treatment, .SDcols = cols_upd ]

```

```{r avg cell type prop, echo =F}
cols2sel <- paste0( c("cells_treatment", sort( names(ct_prop_copy)[grep("avg_ct_", names(ct_prop_copy))] )), collapse = "|") 
dt <- unique(ct_prop_copy[ , .SD, .SDcols= grep( cols2sel, names(ct_prop_copy))])

DT::datatable(dt, rownames = F)
                           
```

## Plots
```{r plot proportions, echo=F}

mdata <- mdata[ cells != "neuron",]
ggplot( mdata, aes(x = cell_type, y = cell_type_prop*100), col= indi, group = treatment ) +
     geom_boxplot(aes( fill= treatment), outlier.shape =  21 ) +
  facet_wrap(~cells) +
   ggtitle("Cell type proportions in the bulk ( deconvolution by Scaden)") +
  theme(legend.position="bottom", legend.box = "vertical")


```
There seems to be a difference in the clusters 0 and 4 treated vs untreated in the cocultures<br>

<b>Is there a significant difference in the cell types between the treated and untreated per cell culture? </b> <br>
<br>
Running a 2-tailed paired sample t-test ( it compares the means between the 2 paired groups of samples) <br> <br>
As the n< 30, need to check if the differences of the pairs follow a normal distribution. Done using the Shapiro-Wilk test ( NULL = data are normally distributed). 


```{r checking for significant difference in the cell type proportions, echo =F, include =F}
#running a paired t-test  - compare the means between two related groups of samples
#head(ct_prop, 2)
ct_prop <- ct_prop[ cells != "neuron", ]
cells_culture <- unique(ct_prop$cells)
cell_types <- names(ct_prop)[grep ("celltype_", names(ct_prop))]
ret_cells <- lapply(cells_culture, function(cel){
  #cel <- cells_culture[1] 
  l_cells <- ct_prop[ cells == cel,  ]
  #head(l_cells, 2)
  ret_celltypes <- lapply(cell_types, function(ct){
    #ct <- cell_types[1]
    cols2sel <-paste0( c(ct, "cells", "treatment", "indi"), collapse="|")
    l_recs <- l_cells[ , .SD, .SDcols= grep( cols2sel, names(l_cells))]
    ddata <- dcast.data.table(l_recs, cells+indi ~treatment, value.var = ct)
    #Shapiro-Wilk normality test for the differences
    shap_wilk_pval <- shapiro.test(ddata$treated - ddata$untreated)$p.value

    pval <- t.test(ddata$treated, ddata$untreated, paired = TRUE, alternative = "two.sided")$p.value
    
    return( data.table( celltype = ct, pval_pairedttest = pval, pval_shapwilk = shap_wilk_pval))
})
  
  ret_celltypes <- rbindlist(ret_celltypes)
  ret_celltypes$cells <- cel

  return(ret_celltypes)
})


ret_cells <- rbindlist(ret_cells)
ret_cells[, fdr := p.adjust(pval_pairedttest, method = "BH"), by=cells] 

cols_upd <- c( "pval_pairedttest", "pval_shapwilk", "fdr")
ret_cells[, (cols_upd) := lapply(.SD, function(x) 
        format(x, scientific = T, digits = 2 )), .SDcols = cols_upd]

dres <- dcast.data.table(ret_cells, celltype~cells, value.var = c( "pval_pairedttest", "fdr", "pval_shapwilk"))

```


<br><br> <b>Shapiro-Wilk normality test p-values</b><br>

pvalues > 0.05 significance level, implying the distribution of the differences are not significantly different from normal distribution. <br>

```{r display shapwilk pvals, echo = F}
cols2sel <-paste0( c("celltype", names(dres)[grep("shapwilk", names(dres))]), collapse="|")

DT::datatable(dres[ , .SD, .SDcols= grep( cols2sel, names(dres))], rownames = F )

```

<br>

<br>
<b> Paired t-test p-values and FDR (multiple comparison test performed using Benjamini & Hochberg method, per cell culture) </b> <br>

<u>All data </u>
```{r display paired ttst pvals, echo = F}
cols2sel <-paste0( c("celltype", names(dres)[grep("pairedttest|fdr", names(dres))]), collapse="|")

DT::datatable(dres[ , .SD, .SDcols= grep( cols2sel, names(dres))], rownames = F)

```


<br><br>
<u> Excluding cell types where proportion < 0.01 </u> <br>
Excluding the cell types 5, 6 and 7
```{r excl prop less that 0.0.1, include= F}
ct2exclude<- paste0("celltype_", seq(5,7))

ret_cells <- ret_cells[ !celltype %in% ct2exclude, ]
ret_cells[, fdr := p.adjust(pval_pairedttest, method = "BH"), by=cells] 

cols_upd <- c( "fdr")
ret_cells[, (cols_upd) := lapply(.SD, function(x) 
        format(as.numeric(x), scientific = T, digits = 2 )), .SDcols = cols_upd]

dres <- dcast.data.table(ret_cells, celltype~cells, value.var = c( "pval_pairedttest", "fdr", "pval_shapwilk"))


```
```{r display excl celltype 5 6 7 paired ttst pvals, echo = F}
cols2sel <-paste0( c("celltype", names(dres)[grep("pairedttest|fdr", names(dres))]), collapse="|")

DT::datatable(dres[ , .SD, .SDcols= grep( cols2sel, names(dres))], rownames = F)

```

