---
title: 'Overlap of genes with increased/decreased editing with the DE and DS genes'
author: "Karishma D'Sa"
date: "17/10/2022"
output:
  html_document:
    self_contained: no
    df_print: paged
    toc: yes
    toc_float: yes
    toc_depth: 6
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
    toc_depth: 6
---
<style type="text/css">
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>
<br><br>

The final list of edited genes provided by AW - <br> 
-	sig editing at the site each site assigned using the  Zscore ( from JACUSA) cut-off 1.96 => equiv to pval 0.05 <br>
-	sites in > 1 gene at the moment <br>


```{r libraries, include = FALSE}

#UpSetR to generate intersections oflists
library(UpSetR)

library(data.table)
library(ggplot2)
library(knitr)
library(ggpubr) 
library(gridExtra)
library(here)
library(tidyverse)
library(magrittr)


source(here("./R/plot_setting.R"))
source(here("./R/common_upsetR.R"))

deseq_rda_path <-here("../results/dge_postdeconv/correct_PC234/") 
outPath <- here("../results/mRNA_editing/dge_ds_overlap/")

if(!dir.exists(outPath)){
  dir.create(outPath, recursive = T)
}
out_res_clustProf <- paste0(outPath, "cluster_profiler/" )

#outPath_dge_go <- paste0(out_res_clustProf, "dge/go/")
outPath_ds_go <- paste0(out_res_clustProf, "ds/go/")

comparison_astr <- "astrocyte_treated_vs_untreated"
comparison_cocul <- "coculture_treated_vs_untreated"
```

```{r loading the editing data, eval=T, echo=F, include=T}
#RNA editing 
ed_data <- fread(here("../files/editing/editing_summary_results_genes.csv"))
head(ed_data, 2)
#names(ed_data)

#excluding the neurons
ed_data <- ed_data[ Sample != "Neuron", ]
dt_genes_edited <- unique(ed_data[, .(Sample, Gene, SYMBOL)]) #irrespective of inc/dec editing and consequence ( as as gene can have > 1 editing sites with different consequence), includes the new edits
table(dt_genes_edited$Sample)
length(unique(dt_genes_edited[ Sample=="Astrocyte", Gene])) 

#editing bg genes
bg_ed <- fread(here("../files/editing/summary_results_all_dif_genelist_background_universe.csv"))
#head(bg_ed)
setnames(bg_ed, old = "SYMBOL", new = "genes")
bg_ed$studied_edit <- "T"
```


```{r loading the DS data, eval=T, echo=T, include=F}
#DS
###
abs_dpsi_cutoff <- 0.1

load(here("../results/leafcutter_additionalFilters_postdeconv/diff_splicing/combGrp_clust_effSize.rda"))
rm(cluster_recs, intron_eff_recs)
comp <- unique(cluster_recs_single_gene_intron$comparison)
cluster_recs_single_gene_intron <- cluster_recs_single_gene_intron[ status == "Success" & !(is.na(genes)),] #319199

cluster_recs_single_gene_intron[, abs_deltapsi := abs(cluster_recs_single_gene_intron$deltapsi)]

ds_genes <- lapply(comp, function(cmp){
    #cmp <- comp[1]
    #head(cluster_recs_single_gene_intron, 2)  
    recs <- cluster_recs_single_gene_intron[ comparison== cmp,]
    #head(recs, 2)
    #select genes of interest - same filter as used to select query gene
    recs <- recs[ p.adjust < 0.05 & abs_deltapsi >= 0.1 & !is.na(genes), ]
    #recs <- recs[ p.adjust < 0.05 & !is.na(genes), ]
    #head(recs[ p.adjust < 0.05 & is.na(genes), ])
    #sigGenes <- unique(recs[ , genes])
    min_recs <- recs[ , .SD[which.min(p.adjust)], by = "genes"] #the 
    #return(unique(min_recs[ , genes]))
    return(min_recs)
})
ds_genes<- rbindlist(ds_genes)
head(ds_genes, 2)
ds_genes[ comparison == "astrocyte_treatment_vs_no_treatment", comparison:= comparison_astr]
ds_genes[ comparison == "astrocyte_neuron_treatment_vs_no_treatment", comparison:= comparison_cocul]
cols2sel <- c(  "genes", "cluster", "p.adjust"   ,      "comparison"  , "deltapsi"   ,  "reg",  "abs_deltapsi")
ds_genes <- ds_genes[ , ..cols2sel ]

bg_ds <-  unique(cluster_recs_single_gene_intron[ , .(genes, comparison)])
bg_ds[ comparison == "astrocyte_treatment_vs_no_treatment", comparison:= comparison_astr]
bg_ds[ comparison == "astrocyte_neuron_treatment_vs_no_treatment", comparison:= comparison_cocul]

print("nBg genes")
table(bg_ds$comparison)

print("DS at FDR < 5% and abs_dpsi >=0.1")
table(ds_genes$comparison)

```



```{r ds and ed, eval=F, include =F}
#genes studied in both
dt_genes_edited[ Sample == "Astrocyte", comparison := comparison_astr ]
dt_genes_edited[ Sample == "Coculture", comparison :=comparison_cocul ]
setnames(dt_genes_edited, old = "SYMBOL", new = "genes")
dt_genes_edited$edit_sig <- "T"    #6012
table( dt_genes_edited$comparison) 

dt_genes_edited <- unique(dt_genes_edited[ , .(genes, comparison, edit_sig)]) #6010 # keeping 1 rec where genes with different ids
bg_ed <- unique(bg_ed[ , .(genes, studied_edit)])

ds_genes$ds_sig <- "T"
dt_ds <- merge(bg_ds, ds_genes, by = c("comparison", "genes"), all=T)
dt_ds$studied_ds <- "T"

ds_ed_all <- merge(dt_ds, dt_genes_edited, by = c("comparison", "genes"), all.x=T) # not considering the genes that were sig edited but not studied for splicing

ds_ed_all$studied_edit <- "F"

ds_ed_all[ genes %in% bg_ed$genes, studied_edit := "T"] # wesill only indicate the geens in DGE not studied in editing but not the genes studied in editing that werent in dge

table(ds_ed_all$studied_edit , ds_ed_all$comparison, exclude=NULL)
   

table(ds_ed_all$studied_ds, ds_ed_all$comparison, exclude = NULL)

ds_ed_all$studied_in <- "ds"             
ds_ed_all[ studied_edit == "T", studied_in := "both" ]  
table(ds_ed_all$studied_in, ds_ed_all$comparison, exclude = NULL)

head( ds_ed_all)
head(ds_ed_all[ ds_sig == "T" & (edit_sig=="F" | is.na(edit_sig)), ])
ds_ed_all[ ds_sig == "T" & edit_sig =="T", sig_in := "both" ]
ds_ed_all[ ds_sig == "T" & (edit_sig=="F" | is.na(edit_sig)), sig_in := "ds" ]
ds_ed_all[ (ds_sig=="F" | is.na(ds_sig)) & edit_sig=="T", sig_in := "edit" ]
table(ds_ed_all$comparison, ds_ed_all$sig_in, exclude=NULL)

```

```{r saving , eval=F, include = F}
fwrite(ds_ed_all, paste0(outPath, "ds_edit_comb_geneset.tsv"), sep="\t", quote=F, row.names = F)

```


```{r ds ed, eval=F, include=F}

source("~/ipscAstrocyteNeuron/ipscAstroNeu/R/common_dge.R") #to get the gene ids for the genes
ds_ed_all <- fread( paste0(outPath, "ds_edit_comb_geneset.tsv"))
ds_ed_all <- replace(ds_ed_all, ds_ed_all=='', NA)
table(ds_ed_all$comparison, ds_ed_all$studied_in, exclude=NULL)

ds_ed_common <- ds_ed_all[ studied_in == "both", ]
  
table(ds_ed_common$comparison, ds_ed_common$sig_in, exclude=NULL)
table(ds_ed_common$comparison, exclude= NULL)
```

```{r saving the ds ed, eval=F, include = F}
fwrite(ds_ed_common, paste0(outPath, "ds_edit_common_geneset.tsv"), sep="\t", quote=F, row.names = F)
```

```{r counts common to the ds and ed, eval=F, include =F}
head(ds_ed_common, 2)
cts_commonset <-  as.data.table(table(ds_ed_common$comparison, ds_ed_common$sig_in, exclude=NULL))
head(cts_commonset, 2)
setnames(cts_commonset , old =c("V1", "V2", "N"), new= c("comparison", "sig_in", "nSig"))
cts_commonset[, total_studied_in_both :=sum(nSig), by=comparison]
cts_commonset[ ,pctSig := (nSig/total_studied_in_both)*100]
cts_commonset[ is.na(sig_in), sig_in:= "neither"]
fwrite(cts_commonset, paste0(outPath, "ds_edit_common_geneset_summ_counts.tsv"), sep="\t", quote=F, row.names = F)

```


#DS, DEd overlap

## Common gene set

<b>Astrocytes - treated vs untreated </b> <br>
```{r summ table astr, eval=T, echo=F}
dt_summ <- fread(paste0(outPath, "ds_edit_common_geneset_summ_counts.tsv"))
DT::datatable(dt_summ[ comparison== "astrocyte_treated_vs_untreated", ])
```
<br>

<br>
<b>Co-culture - treated vs untreated </b><br>
```{r summ table co-cul, eval=T, echo=F}
DT::datatable(dt_summ[ comparison== "coculture_treated_vs_untreated", ])
```

<b>Fisher's Exact test run </b><br>
```{r stat test for overlap between the DGE and DEd in the common geneset, eval =T, echo=F}
ds_ed_all <- fread( paste0(outPath, "ds_edit_comb_geneset.tsv"))
ds_ed_all <- replace(ds_ed_all, ds_ed_all=='', NA)
common <- ds_ed_all[ studied_in== "both", ]
table(common$ds_sig, exclude=NULL)

common[ is.na(ds_sig),  ds_sig := "ds_sig_F"]
common[ ds_sig == "T",  ds_sig := "ds_sig_T"]
comp <- unique(common$comparison)
head(common)
ret_vals <- lapply(comp, function(cmp){
  cat("\n=======\n", cmp, "\n")
  df_conting <- as.data.frame.matrix(table(common[ comparison ==cmp , .(edit_sig, ds_sig) ], exclude=NULL))
  print(df_conting)
  fisher <- fisher.test(df_conting)$p.value
  cat("\nFisher's Exact Test p-value =",fisher)
  return(T)
  })

```

###Enrichment analysis 
Based on the decision to go with the bg of all sig DGE <br>
- of the differentially expressed genes, what's special about the ones that are edited <br><br>

running the same with the DS - taking the genes studied in both DS and ed <br>
background : all DS and  query : sigDS and editing <br><br>
 No enrichment found in any of gene cluster
 
